---
title: "Asymmetric functional prediction bands for assessing the validity of biomechanical curve data"
author:
  - name: Daniel Koska
    email: daniel.koska@hsw.tu-chemnitz.de
    affiliation: Chemnitz University of Technology, Research Methodology and Data Analysis in Biomechanics
    footnote: Corresponding Author
  - name: Doris Oriwol
    email: doris.oriwol@partner.kit.edu
    affiliation: Karlsruhe Institute of Technology
  - name: Christian Maiwald
    email: christian.maiwal@hsw.tu-chemnitz.de
    affiliation: Chemnitz University of Technology, Research Methodology and Data Analysis in Biomechanics
# address:
#   - code: Chemnitz University of Technology
#     address:  Thüringer Weg 11, 09126 Chemnitz
#   - code: Another University
#     address: Department, Street, City, State, Zip
abstract: |
  This is the abstract.

date: "`r format(Sys.time(), '%d %B, %Y')`"
journal: "An awesome journal"
# bibliography: sendaFAB.bib
# output: rticles::elsevier_article
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
---

### TODO
- bar

```{r echo = FALSE, warning = FALSE, message = FALSE}
library(ggplot2)

dir.script <- "C:/Users/Daniel/Desktop/tmp/floa/R"
dir.data <- "C:/Users/Daniel/Desktop/tmp/floa/R/examples"

# setwd(dir.script)
source(paste0(dir.script, "/example_data.R"))
source(paste0(dir.script, "/draw_clusters.R"))
source(paste0(dir.script, "/floa_rcb.R"))
source(paste0(dir.script, "/floa_point.R"))
source(paste0(dir.script, "/plot_loa.R"))
source(paste0(dir.script, "/get_coverage.R"))
source(paste0(dir.script, "/crossval_coverage.R"))
source(paste0(dir.script, "/plot_cov_ver.R"))
# source("fdaDelta.R")
```

# Introduction

# Methods

## Limits of agreement methods

FLoA derived by different methods are compared

* Randomized Cluster Bootstrap (floa.rcb)
  + Different strategies

* Point-by-point Gaussian intervals (floa.point)

```{r echo = FALSE, warning = FALSE, message = FALSE}
# Wrapper function for example data sets. Function arguments:
#
# (* Empirical validation data: "imu_mc")
# * Smooth, wave data (normal error, constant variance, no trend): "smooth"
# * Smooth wave data with nonlinear trend (constant variance): "smooth_trend"
# * Data with non-gaussian (Weibull distributed) error (no trend): "non_gaussian"
# * Data with shock peaks (no bias, no trend): "shock"

data <- example_data(dat = "smooth", dir.data)

n.boot <- 100

floa <- floa_rcb(data, n.boot, ver = "v2")

# Mean and SD are calculated across all strides (and subjects).
# No bootstrap or other resampling strategies are applied.
floa.point <- floa_point(data)
```

## Validation

### Synthetic data sets

```{r echo = FALSE, warning = FALSE, message = FALSE}

```

### Smooth, wave data (normal error, constant variance, no trend)

```{r echo = FALSE, warning = FALSE, message = FALSE}
data <- example_data(dat = "smooth", dir.data)

data.single.mc <- subset(data, device == "MC")
data.single.imu <- subset(data, device == "IMU")

PLOT <- ggplot(data = data.single.mc, aes(x = frame, y = value, group = strideID)) + # , colour = subjectID
  geom_line(alpha = 0.7) +
  geom_line(data = data.single.imu, aes(x = frame, y = value, group = strideID, col = "red"), alpha = 0.3) + #, colour = subjectID
  labs(x = "Time-normalized signal [%]", y = "Difference") +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 20),
        axis.title.x = element_text(size = 22),
        axis.text.y = element_text(size = 20),
        axis.title.y = element_text(size = 22),
        legend.position = "none")

PLOT
```

### Smooth wave data with nonlinear trend

```{r echo = FALSE, warning = FALSE, message = FALSE}
data <- example_data(dat = "smooth_trend", dir.data)

data.single.mc <- subset(data, device == "MC")
data.single.imu <- subset(data, device == "IMU")

PLOT <- ggplot(data = data.single.mc, aes(x = frame, y = value, group = strideID)) + # , colour = subjectID
  geom_line(alpha = 0.7) +
  geom_line(data = data.single.imu, aes(x = frame, y = value, group = strideID, col = "red"), alpha = 0.3) + #, colour = subjectID
  labs(x = "Time-normalized signal [%]", y = "Difference") +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 20),
        axis.title.x = element_text(size = 22),
        axis.text.y = element_text(size = 20),
        axis.title.y = element_text(size = 22),
        legend.position = "none")

PLOT
```

#### Data with non-gaussian (Weibull distributed) error (no trend)

```{r echo = FALSE, warning = FALSE, message = FALSE}
data <- example_data(dat = "non_gaussian", dir.data)

data.single.mc <- subset(data, device == "MC")
data.single.imu <- subset(data, device == "IMU")

PLOT <- ggplot(data = data.single.mc, aes(x = frame, y = value, group = strideID)) + # , colour = subjectID
  geom_line(alpha = 0.7) +
  geom_line(data = data.single.imu, aes(x = frame, y = value, group = strideID, col = "red"), alpha = 0.3) + #, colour = subjectID
  labs(x = "Time-normalized signal [%]", y = "Difference") +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 20),
        axis.title.x = element_text(size = 22),
        axis.text.y = element_text(size = 20),
        axis.title.y = element_text(size = 22),
        legend.position = "none")

PLOT
```

### Cross validation

Leave-one (subject) out method to estimate the achieved coverage

Versions:

* v1   : n = length(subjects) random strides from all strides

* v1.1 : Functional data version of v1

* v2   : One random stride per subject

* v3   : Fetch a SINGLE random stride from all strides

<!-- $$ -->
<!-- RMSE_{delta} = \sqrt{\frac{1}{N}(target_1 - pedal_1)^2 + (target_2 - pedal_2)^2 + ... + (target_N - pedal_N)^2} -->
<!-- $$ -->

<!-- $$ -->
<!-- RMSE_{pedal} = \sqrt{\frac{1}{n}[(pedal_1 - \mu)^2 + (m_2 - \mu)^2 + ... + (pedal_N - \mu)^2]}, where \mu = \frac{1}{N}(pedal_1 + ... + pedal_N) -->
<!-- $$ -->

<!-- $$ -->
<!-- Z = \{z_1, z_2, z_3,...,z_i\} -->
<!-- $$ -->

# Results

### Cross validation

#### v1 (n = length(subjects) random strides from all strides)

```{r echo = FALSE, warning = FALSE, message = FALSE}
data <- example_data(dat = "smooth", dir.data)
```

```{r echo = FALSE, warning = FALSE, message = FALSE}
cover.cross.v1 <- crossval_coverage(data, n.boot, method = "all", ver = "v1")

plot_cov_ver(cover.cross.v1)
```

### V2 (One random stride per subject)

```{r echo = FALSE, warning = FALSE, message = FALSE}
cover.cross.v2 <- crossval_coverage(data, n.boot, method = "all", ver = "v2")

plot_cov_ver(cover.cross.v2)
```

### v3 (Fetch a SINGLE random stride from all strides)

```{r echo = FALSE, warning = FALSE, message = FALSE}
cover.cross.v3 <- crossval_coverage(data, n.boot, method = "all", ver = "v3")

plot_cov_ver(cover.cross.v3)
```

# Discussion


# Conclusion

